My title

### 2024-03-29 paper/paper.qmd
added the data section

---
title: "My title"
subtitle: "My subtitle if needed"
author: Ahnaf Alam
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
toc: true
---

@@

data |>
  ggplot(aes(x = date, y = food_expenditure)) +
  geom_line() +
  theme_minimal() +
  labs(x = "Date",
       y = "Real food expenditure",
       )

data |>
  ggplot(aes(x = date, y =income_expenditure)) +
  geom_line() +
  theme_minimal() +
  labs(x = "Date",
       y = "Real disposable income",
       )
```

@@

modelsummary::modelsummary(
  list(
    "Simple linear" = first_model,
    "Multiple linear" = second_model
  ),
  statistic = "mad",
  fmt = 2
)
```

@@

Our results are summarized in @tbl-modelresults.

>>>>

---
title: "My title"
subtitle: "My subtitle if needed"
author: Ahnaf Alam
thanks: "Code and data are available at https://github.com/AhnafAlam1/income_and_expenditure/tree/main"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
toc: true
---

@@

data |>
  ggplot(aes(x = date, y = food_expenditure)) +
  geom_line() +
  theme_minimal() +
  labs(x = "Date",
       y = "Changes in real food expenditure",
       )

data |>
  ggplot(aes(x = date, y =income_expenditure)) +
  geom_line() +
  theme_minimal() +
  labs(x = "Date",
       y = "Changes in real disposable income",
       )
```

@@

modelsummary::modelsummary(
  list(
    "Model 1" = first_model,
    "Model 2" = second_model
  ),
  statistic = "mad",
  fmt = 2
)
```

@@

Our findings are outlined in @tbl-3. In simple linear regression, we observed an estimated intercept of 24.07 billion USD. Additionally, for every 1 billion USD increase in disposable income, there's a corresponding rise in food expenditure of $70 million USD. The small standard deviation suggests a low level of variability and uncertainty.

In multiple linear regression, we found an estimated intercept of -97.40 billion USD. With each $1 billion USD increase in income, food expenditure increased by 20 million USD. Similar to the simple regression model, the standard deviation is negligibly small.

Both models exhibit similar adjusted $R^2$ values, yet the second model slightly outperforms the first with a higher value (0.993 vs 0.990), indicating a superior fit. Notably, the root mean square error (RSME) for the second model is substantially lower than that of the first (10.36 vs 27.49), indicating enhanced predictive accuracy.

@fig-1 illustrates quarterly changes in disposable income and food expenditure from 2007 to 2022. Both variables show consistent growth trends, with 33.9% increase in disposable income and 32.5% increase in expenditure in food in this period. We see a notable surge in disposable income in 2020 due to COVID-era relief efforts. However, this increase was followed by a decline as restrictions eased, allowing for increased spending towards the end of 2021.

@fig-2 displays the distribution of expenditure across categories, highlighting durable and non-durable goods as the largest expenditures. This aligns with expectations given the nature of these commodities. Following closely are food expenditure and disposable income, reflecting a strong correlation evident in @fig-3, where a clear linear relationship is observed with small standard deviations. Finally, healthcare and services expenditure are clustered together.
